### 2020-02-11 submission/empirical2_TRfmt.tex
fixed bib refs

%\documentclass[12pt]{article}
\documentclass[11pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
%\addtolength{\topmargin}{-0.5in}
\addtolength{\topmargin}{-1.0in}
\addtolength{\textheight}{1.7in}
\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{framed}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
%%% \usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{caption}
%%\DeclareCaptionType{copyrightbox}
%\usepackage{subcaption}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{color}
\usepackage{xcolor}
%%\usepackage{enumitem}
\usepackage{paralist}

%\usepackage{subfig}
\usepackage{subfigure}

\usepackage{longtable} 
\usepackage{makecell}

\usepackage[multiple]{footmisc}
\usepackage[section]{placeins}

\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     linkcolor    = blue,
     citecolor    = green
}

%-----------------------------------------------------------------------

\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\comment}[1]{\textcolor{blue}{#1}}
\newcommand{\awk}[1]{\textcolor{green}{#1}}

\newcommand{\argmin}{\text{argmin}}
\newcommand{\Probab}[1]{\mbox{}{\bf{Pr}}\left[#1\right]}
\newcommand{\Expect}[1]{\mbox{}{\bf{E}}\left[#1\right]}
\newcommand{\ExpectBracket}[1]{\mbox{}\langle#1\rangle}

% Here are two macros for comments.
\newcommand {\nred}[1]{{\color{red}\sf{[#1]}}}
\newcommand {\ngreen}[1]{{\color{green}\sf{[#1]}}}
\newcommand {\ncyan}[1]{{\color{cyan}\sf{[#1]}}}
\newcommand {\michael}[1]{{\color{red}\sf{[michael: #1]}}}
\newcommand {\charles}[1]{{\color{blue}\sf{[charles: #1]}}}
\newcommand {\serena}[1]{{\color{orange}\sf{[charles: #1]}}}

\usepackage[normalem]{ulem}

%-----------------------------------------------------------------------

\begin{document}

\title{%
Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data
}

\author{%
Charles H. Martin\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{charles@CalculationConsulting.com}.} 
\and 
Tongsu (Serena) Peng\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{serenapeng7@gmail.com}.}
\and
Michael W. Mahoney\thanks{ICSI and Department of Statistics, University of California at Berkeley, Berkeley, CA 94720, \texttt{mmahoney@stat.berkeley.edu}.}
}

\date{}
\maketitle

\begin{abstract}
\input{abstract_file}
\end{abstract}

%GLG | Powered by Box

\input{introduction}
\input{background}
\input{methods}
\input{cv_models}
\input{nlp_models}
\input{all_models}
\input{conclusion}

\noindent
\paragraph{Acknowledgements.}
MWM would like to acknowledge ARO, DARPA, NSF, and ONR for providing partial support of this work.
We would also like to thank Amir Khosrowshahi and colleagues at Intel for helpful discussion regarding the Group Regularization distillation technique.

\bibliographystyle{unsrt}
%\bibliographystyle{plain}
{\small
%\bibliography{gen_gap}
%\bibliography{dnns}
\bibliography{dnns,gen_gap}
}

\appendix
\input{appendix}

\end{document}