### 2020-02-12 submission/conclusion.tex

\section{Conclusion}
\label{sxn:conc}

We have developed (based on strong theory) and evaluated (on a large corpus of publicly-available pretrained models from CV and NLP) methods to predict trends in the quality of state-of-the-art neural networks---without access to training or testing data.
Prior to our work, it was not obvious that norm-based metrics would perform well to predict trends in quality \emph{across} models (as they are usually used \emph{within} a given model or parameterized model class, e.g., to bound generalization error or to construct regularizers).
Our results are the first to demonstrate that they can be used for this important practical problem.
That PL-based metrics perform better (than norm-based metrics) should not be surprising---at least to those familiar with the statisical mechanics of heavy tailed and strongly correlated systems~\cite{BouchaudPotters03, SornetteBook, BP11, bun2017} (since our use of PL exponents is designed to capture the idea that well-trained models capture correlations over many size scales in the data).
Again, though, our results are the first to demonstrate this.
It is also gratifying that this approach can be used to provide fine-scale insight (such as rationalizing the flow of correlations or the collapse of size scale) throughout a network.

\ncyan{
We conclude with a few thoughts on what a \emph{practical theory} of DNNs should look like.

To do so, we distinguish between two types of theories:
what may be called an \emph{non-empirical or analogical theory}, in which one creates a very simple model that can be analyzed rigorously, and one argues that the model has properties that are analogous the the system of interest; and 
what may be called a \emph{semi-empirical theory}, in which there exists a rigorous asymptotic theory that has asymptotic parameters, and one adjusts or fits the parameters to the finite data.

often one simple enough to perform ``rigorous'' analysis, and one argues that although simple

%\michael{MM TO DO.}
%%
%We distinguish between what we will call a
%\emph{phenomenological theory}
%(that describes empirical relationship of phenomena to each other, in a way which is consistent with fundamental theory, but is not directly derived from that theory)
%and what can be called a 
%\emph{first principles theory} 
%(that is applicable to toy models, but that does not scale up to realistic systems if one includes realistic apsects of realistic systems).
%%
%For most complex highly-engineered systems (aside from complex AI/ML systems), one \emph{uses} phenomenological theory rather than first principles theory. 
%(One does not try to solve the Schr\"odinger equation if one is interested in building a bridge or flying an airplane.)
%%

Much of the work on the theory of DNNs focuses on the former type of theory, but our approach is of the latter type.
Our results, however, which are based on our \emph{use} of sophisticated statistical mechanics theory to solve an important practical DNN problems, suggests that our approach should be of interest more generally for those interested in developing a practical DNN theory.
}