### 2020-10-02 submission/empirical2_NatComfmt.tex
more on address presubmission comments

%\documentclass[12pt]{article}
\documentclass[11pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
%\addtolength{\topmargin}{-0.5in}
\addtolength{\topmargin}{-1.0in}
\addtolength{\textheight}{1.7in}
\newlength{\defbaselineskip}
\setlength{\defbaselineskip}{\baselineskip}

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{framed}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
%%% \usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{caption}
%%\DeclareCaptionType{copyrightbox}
%\usepackage{subcaption}
\usepackage{url}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{color}
\usepackage{xcolor}
%%\usepackage{enumitem}
\usepackage{paralist}

%\usepackage{subfig}
\usepackage{subfigure}

\usepackage{longtable} 
\usepackage{makecell}

\usepackage[multiple]{footmisc}
\usepackage[section]{placeins}

\usepackage{hyperref}
\hypersetup{
     colorlinks   = true,
     linkcolor    = blue,
     citecolor    = green
}

%-----------------------------------------------------------------------

\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\comment}[1]{\textcolor{blue}{#1}}
\newcommand{\awk}[1]{\textcolor{green}{#1}}

\newcommand{\argmin}{\text{argmin}}
\newcommand{\Probab}[1]{\mbox{}{\bf{Pr}}\left[#1\right]}
\newcommand{\Expect}[1]{\mbox{}{\bf{E}}\left[#1\right]}
\newcommand{\ExpectBracket}[1]{\mbox{}\langle#1\rangle}

% Here are two macros for comments.
\newcommand {\nred}[1]{{\color{red}\sf{[#1]}}}
\newcommand {\ngreen}[1]{{\color{green}\sf{[#1]}}}
\newcommand {\ncyan}[1]{{\color{cyan}\sf{[#1]}}}

%% MWM: Do the following to remove the comments from the pdf.
\newcommand {\michael}[1]{{ }}
%\newcommand {\michael}[1]{{\color{red}\sf{[michael: #1]}}}

\newcommand {\charles}[1]{{\color{blue}\sf{[charles: #1]}}}
\newcommand {\serena}[1]{{\color{orange}\sf{[charles: #1]}}}

\usepackage[normalem]{ulem}

%-----------------------------------------------------------------------

\begin{document}

\title{%
Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data
}

\author{%
Charles H. Martin\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{charles@CalculationConsulting.com}.} 
\and 
Tongsu (Serena) Peng\thanks{Calculation Consulting, 8 Locksley Ave, 6B, San Francisco, CA 94122, \texttt{serenapeng7@gmail.com}.}
\and
Michael W. Mahoney\thanks{ICSI and Department of Statistics, University of California at Berkeley, Berkeley, CA 94720, \texttt{mmahoney@stat.berkeley.edu}.}
\thanks{Corresponding author.}
}

\date{}
\maketitle

\begin{abstract}
\input{NatCom_abstract_file}
\end{abstract}

%GLG | Powered by Box

\input{NatCom_introduction}

\section{Results}
After describing our overall approach, we study in detail three well-known CV architectures (the VGG, ResNet, and DenseNet series of models).
Then, we look in detail at several variations of a popular NLP architecture (the OpenAI GPT and GPT2 models), and we present results from a broader analysis of hundreds of pretrained DNN~models.

\input{NatCom_overall_approach}
\input{NatCom_cv_models}
\input{NatCom_nlp_models}
\input{NatCom_all_models}

\input{NatCom_discussion}

\input{NatCom_methods}

\noindent
\paragraph{Acknowledgements.}
MWM would like to acknowledge ARO, DARPA, NSF, and ONR as well as the UC Berkeley BDD project and a gift from Intel for providing partial support of this work.
Our conclusions do not necessarily reflect the position or the policy of our sponsors, and no official endorsement should be inferred.
We would also like to thank Amir Khosrowshahi and colleagues at Intel for helpful discussion regarding the Group Regularization distillation technique.

\bibliographystyle{unsrt}
%\bibliographystyle{plain}
{\small
%\bibliography{gen_gap}
\bibliography{dnns}
%\bibliography{dnns,gen_gap}
}

\appendix
\input{NatCom_appendix}

\end{document}