### 2021-01-01 submission/figs_largeScale_A.tex
put all those figs in appendix

\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_lognorm}
        \label{fig:summary_regressions_A_01}
    }
    \subfigure[ imagenet-1k--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_spectralnormlog}
        \label{fig:summary_regressions_A_02}
    }
    \subfigure[ imagenet-1k--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_alpha_weighted}
        \label{fig:summary_regressions_A_03}
    }
    \subfigure[ imagenet-1k--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ResNet_logpnorm}
        \label{fig:summary_regressions_A_04}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_lognorm}
        \label{fig:summary_regressions_A_05}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_spectralnormlog}
        \label{fig:summary_regressions_A_06}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_alpha_weighted}
        \label{fig:summary_regressions_A_07}
    }
    \subfigure[ imagenet-1k--EfficientNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_EfficientNet_logpnorm}
        \label{fig:summary_regressions_A_08}
    }
    \subfigure[ imagenet-1k--PreResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_lognorm}
        \label{fig:summary_regressions_A_09}
    }
    \subfigure[ imagenet-1k--PreResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_spectralnormlog}
        \label{fig:summary_regressions_A_10}
    }
    \subfigure[ imagenet-1k--PreResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_alpha_weighted}
        \label{fig:summary_regressions_A_11}
    }
    \subfigure[ imagenet-1k--PreResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_PreResNet_logpnorm}
        \label{fig:summary_regressions_A_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--ResNet; 
             imagenet-1k--EfficientNet; 
             and 
             imagenet-1k--PreResNet; 
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_A}
\end{figure}