### 2021-01-01 submission/figs_largeScale_B.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--ShuffleNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_lognorm}
        \label{fig:summary_regressions_B_01}
    }
    \subfigure[ imagenet-1k--ShuffleNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_spectralnormlog}
        \label{fig:summary_regressions_B_02}
    }
    \subfigure[ imagenet-1k--ShuffleNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_alpha_weighted}
        \label{fig:summary_regressions_B_03}
    }
    \subfigure[ imagenet-1k--ShuffleNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ShuffleNet_logpnorm}
        \label{fig:summary_regressions_B_04}
    }
    \subfigure[ imagenet-1k--VGG, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_lognorm}
        \label{fig:summary_regressions_B_05}
    }
    \subfigure[ imagenet-1k--VGG, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_spectralnormlog}
        \label{fig:summary_regressions_B_06}
    }
    \subfigure[ imagenet-1k--VGG, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_alpha_weighted}
        \label{fig:summary_regressions_B_07}
    }
    \subfigure[ imagenet-1k--VGG, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_VGG_logpnorm}
        \label{fig:summary_regressions_B_08}
    }
    \subfigure[ imagenet-1k--DLA, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_lognorm}
        \label{fig:summary_regressions_B_09}
    }
    \subfigure[ imagenet-1k--DLA, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_spectralnormlog}
        \label{fig:summary_regressions_B_10}
    }
    \subfigure[ imagenet-1k--DLA, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_alpha_weighted}
        \label{fig:summary_regressions_B_11}
    }
    \subfigure[ imagenet-1k--DLA, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DLA_logpnorm}
        \label{fig:summary_regressions_B_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--ShuffleNet;
             imagenet-1k--VGG;
             and
             imagenet-1k--DLA;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_B}
\end{figure}