### 2021-01-01 submission/figs_largeScale_C.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--HRNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_lognorm}
        \label{fig:summary_regressions_C_01}
    }
    \subfigure[ imagenet-1k--HRNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_spectralnormlog}
        \label{fig:summary_regressions_C_02}
    }
    \subfigure[ imagenet-1k--HRNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_alpha_weighted}
        \label{fig:summary_regressions_C_03}
    }
    \subfigure[ imagenet-1k--HRNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_HRNet_logpnorm}
        \label{fig:summary_regressions_C_04}
    }
    \subfigure[ imagenet-1k--DRN-C, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_lognorm}
        \label{fig:summary_regressions_C_05}
    }
    \subfigure[ imagenet-1k--DRN-C, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_spectralnormlog}
        \label{fig:summary_regressions_C_06}
    }
    \subfigure[ imagenet-1k--DRN-C, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_alpha_weighted}
        \label{fig:summary_regressions_C_07}
    }
    \subfigure[ imagenet-1k--DRN-C, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_DRN_C_logpnorm}
        \label{fig:summary_regressions_C_08}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_lognorm}
        \label{fig:summary_regressions_C_09}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_spectralnormlog}
        \label{fig:summary_regressions_C_10}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_alpha_weighted}
        \label{fig:summary_regressions_C_11}
    }
    \subfigure[ imagenet-1k--SqueezeNext, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNext_logpnorm}
        \label{fig:summary_regressions_C_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--HRNet;
             imagenet-1k--DRN-C;
             and
             imagenet-1k--SqueezeNext;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_C}
\end{figure}