### 2021-01-01 submission/figs_largeScale_D.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--ESPNetv2, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_lognorm}
        \label{fig:summary_regressions_D_01}
    }
    \subfigure[ imagenet-1k--ESPNetv2, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_spectralnormlog}
        \label{fig:summary_regressions_D_02}
    }
    \subfigure[ imagenet-1k--ESPNetv2, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_alpha_weighted}
        \label{fig:summary_regressions_D_03}
    }
    \subfigure[ imagenet-1k--ESPNetv2, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ESPNetv2_logpnorm}
        \label{fig:summary_regressions_D_04}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_lognorm}
        \label{fig:summary_regressions_D_05}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_spectralnormlog}
        \label{fig:summary_regressions_D_06}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_alpha_weighted}
        \label{fig:summary_regressions_D_07}
    }
    \subfigure[ imagenet-1k--SqueezeNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_SqueezeNet_logpnorm}
        \label{fig:summary_regressions_D_08}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_lognorm}
        \label{fig:summary_regressions_D_09}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_spectralnormlog}
        \label{fig:summary_regressions_D_10}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_alpha_weighted}
        \label{fig:summary_regressions_D_11}
    }
    \subfigure[ imagenet-1k--ProxylessNAS, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_ProxylessNAS_logpnorm}
        \label{fig:summary_regressions_D_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--ESPNetv2;
             imagenet-1k--SqueezeNet;
             and
             imagenet-1k--ProxylessNAS;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_D}
\end{figure}