### 2021-01-01 submission/figs_largeScale_E.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ imagenet-1k--IGCV3, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_lognorm}
        \label{fig:summary_regressions_E_01}
    }
    \subfigure[ imagenet-1k--IGCV3, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_spectralnormlog}
        \label{fig:summary_regressions_E_02}
    }
    \subfigure[ imagenet-1k--IGCV3, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_alpha_weighted}
        \label{fig:summary_regressions_E_03}
    }
    \subfigure[ imagenet-1k--IGCV3, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_imagenet_1k_IGCV3_logpnorm}
        \label{fig:summary_regressions_E_04}
    }
    \subfigure[ cifar-10--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_lognorm}
        \label{fig:summary_regressions_E_05}
    }
    \subfigure[ cifar-10--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_spectralnormlog}
        \label{fig:summary_regressions_E_06}
    }
    \subfigure[ cifar-10--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_alpha_weighted}
        \label{fig:summary_regressions_E_07}
    }
    \subfigure[ cifar-10--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_ResNet_logpnorm}
        \label{fig:summary_regressions_E_08}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_lognorm}
        \label{fig:summary_regressions_E_09}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_spectralnormlog}
        \label{fig:summary_regressions_E_10}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_alpha_weighted}
        \label{fig:summary_regressions_E_11}
    }
    \subfigure[ cifar-10--DIA-ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_DIA_ResNet_logpnorm}
        \label{fig:summary_regressions_E_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             imagenet-1k--IGCV3;
             cifar-10--ResNet;
             and
             cifar-10--DIA-ResNet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_E}
\end{figure}