### 2021-01-01 submission/figs_largeScale_F.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ cifar-10--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_lognorm}
        \label{fig:summary_regressions_F_01}
    }
    \subfigure[ cifar-10--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_spectralnormlog}
        \label{fig:summary_regressions_F_02}
    }
    \subfigure[ cifar-10--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_alpha_weighted}
        \label{fig:summary_regressions_F_03}
    }
    \subfigure[ cifar-10--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_10_SENet_logpnorm}
        \label{fig:summary_regressions_F_04}
    }
    \subfigure[ cifar-100--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_lognorm}
        \label{fig:summary_regressions_F_05}
    }
    \subfigure[ cifar-100--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_spectralnormlog}
        \label{fig:summary_regressions_F_06}
    }
    \subfigure[ cifar-100--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_alpha_weighted}
        \label{fig:summary_regressions_F_07}
    }
    \subfigure[ cifar-100--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_ResNet_logpnorm}
        \label{fig:summary_regressions_F_08}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_lognorm}
        \label{fig:summary_regressions_F_09}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_spectralnormlog}
        \label{fig:summary_regressions_F_10}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_alpha_weighted}
        \label{fig:summary_regressions_F_11}
    }
    \subfigure[ cifar-100--DIA-ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_DIA_ResNet_logpnorm}
        \label{fig:summary_regressions_F_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             cifar-10--SENet;
             cifar-100--ResNet;
             and
             cifar-100--DIA-ResNet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_F}
\end{figure}