### 2021-01-01 submission/figs_largeScale_G.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ cifar-100--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_lognorm}
        \label{fig:summary_regressions_G_01}
    }
    \subfigure[ cifar-100--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_spectralnormlog}
        \label{fig:summary_regressions_G_02}
    }
    \subfigure[ cifar-100--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_alpha_weighted}
        \label{fig:summary_regressions_G_03}
    }
    \subfigure[ cifar-100--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_SENet_logpnorm}
        \label{fig:summary_regressions_G_04}
    }
    \subfigure[ cifar-100--WRN, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_lognorm}
        \label{fig:summary_regressions_G_05}
    }
    \subfigure[ cifar-100--WRN, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_spectralnormlog}
        \label{fig:summary_regressions_G_06}
    }
    \subfigure[ cifar-100--WRN, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_alpha_weighted}
        \label{fig:summary_regressions_G_07}
    }
    \subfigure[ cifar-100--WRN, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_cifar_100_WRN_logpnorm}
        \label{fig:summary_regressions_G_08}
    }
    \subfigure[ svhn--ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_lognorm}
        \label{fig:summary_regressions_G_09}
    }
    \subfigure[ svhn--ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_spectralnormlog}
        \label{fig:summary_regressions_G_10}
    }
    \subfigure[ svhn--ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_alpha_weighted}
        \label{fig:summary_regressions_G_11}
    }
    \subfigure[ svhn--ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_ResNet_logpnorm}
        \label{fig:summary_regressions_G_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             cifar-100--SENet;
             cifar-100--WRN;
             and
             svhn--ResNet;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_G}
\end{figure}