### 2021-01-01 submission/figs_largeScale_H.tex

 
\begin{figure}[t]
    \centering
    \subfigure[ svhn--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_lognorm}
        \label{fig:summary_regressions_H_01}
    }
    \subfigure[ svhn--DIA-ResNet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_spectralnormlog}
        \label{fig:summary_regressions_H_02}
    }
    \subfigure[ svhn--DIA-ResNet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_alpha_weighted}
        \label{fig:summary_regressions_H_03}
    }
    \subfigure[ svhn--DIA-ResNet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_DIA_ResNet_logpnorm}
        \label{fig:summary_regressions_H_04}
    }
    \subfigure[ svhn--SENet, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_lognorm}
        \label{fig:summary_regressions_H_05}
    }
    \subfigure[ svhn--SENet, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_spectralnormlog}
        \label{fig:summary_regressions_H_06}
    }
    \subfigure[ svhn--SENet, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_alpha_weighted}
        \label{fig:summary_regressions_H_07}
    }
    \subfigure[ svhn--SENet, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_SENet_logpnorm}
        \label{fig:summary_regressions_H_08}
    }
    \subfigure[ svhn--WRN, $\langle\log\Vert\cdot\Vert_{F}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_lognorm}
        \label{fig:summary_regressions_H_09}
    }
    \subfigure[ svhn--WRN, $\langle\log\Vert\cdot\Vert_{\infty}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_spectralnormlog}
        \label{fig:summary_regressions_H_10}
    }
    \subfigure[ svhn--WRN, $\hat{\alpha}$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_alpha_weighted}
        \label{fig:summary_regressions_H_11}
    }
    \subfigure[ svhn--WRN, $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$ ]{
        \includegraphics[width=3.7cm]{img/omsr_svhn_WRN_logpnorm}
        \label{fig:summary_regressions_H_12}
    }
    \caption{Regression plots for model-dataset pairs, based on data from Table~\ref{table:MSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:KTresults}.
             Each row corresponds to a different dataset-model pair:
             svhn--DIA-ResNet;
             svhn--SENet;
             and
             svhn--WRN;
             repsectively.
             Each column corresponds to a different metric:
             $\langle\log\Vert\cdot\Vert_{F}\rangle$; 
             $\langle\log\Vert\cdot\Vert_{\infty}\rangle$; 
             $\hat{\alpha}$; 
             and
             $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$;
             repsectively.
            }
    \label{fig:summary_regressions_H}
\end{figure}