### 2021-01-02 submission/table_3.tex

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|p{1.25in}|c|c|c|c|}
\hline
  & $\log\Vert\cdot\Vert^{2}_{F}$ & $\log\Vert\cdot\Vert^{2}_{\infty}$ & $\hat{\alpha}$ & $\log\Vert\cdot\Vert^{\alpha}_{\alpha}$ \\
\hline

RMSE (mean) & 4.84 & 5.56 & 4.58& 4.54 \\
RMSE (std) & 9.14 & 9.17 & 9.16& 9.17 \\
\hline
R2 (mean) & 3.9 & 3.85 & 3.89& 3.89 \\
R2 (std) & 9.34 & 9.36 & 9.34& 9.34 \\
\hline
R2 adjusted (mean) & 3.82 & 3.77 & 3.82& 3.83 \\
R2 adjusted (std) & 9.38 & 9.4 & 9.37& 9.37 \\
\hline
Kendal-tau (mean) & 3.84 & 3.77 & 3.85& 3.85 \\
Kendal-tau (std) & 9.37 & 9.4 & 9.36& 9.36 \\
\hline

\hline
\end{tabular}
\end{center}
\caption{Comparison of linear regression fits for different average Log Norm and Weighted Alpha metrics across 5 CV datasets, 17 architectures, covering \
108 (out of over 400) different pretrained DNNs.
         We include regressions only for architectures with five or more data points, and which are positively correlated with test error.
         These results can be readily reproduced using the Google Colab notebooks.
         (See the Supplementary Information
         %~\ref{sxn:appendix}                                                                                                                             
         for details.)}
\label{table:results}
\end{table}