### 2021-01-11 submission/NatCom_all_models.tex
smooth out a bit

\subsection{Comparing Hundreds of Models}
\label{sxn:all_cv_models}

%%\michael{Refine text in this subsection after the appendix is finished.  Things to emphasize: we have a huge amount of data that we are summarizing very succicently; the table here is the tip of the iceberg; the results show our metrics are good, as a sanity check; they show that you can get some results in general, but looking at the details in the appendix show that results are more reliable within a given architecture series than across different architectures; point to out publicly-availab e repo; mention that we do Kendall-$\tau$ rank correlation also.}

We have performed a large-scale analysis of hundreds of publicly-available models.
This broader analysis is on a much larger set of CV and NLP models, with a more diverse set of architectures, that have been developed for a wider range of tasks; and it complements the previous more detailed analysis on CV and NLP models, where we have analyzed only a single architecture series at a time.
See the Supplementary Information
%~\ref{sxn:appendix} 
(and our publicly-available repo)
for details.
To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares 
%(similar results are seen with rank correlation metrics) 
to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables), and we report the RMSE, the $R^2$ (R2) regresssion metric, and the Kendal-$\tau$ rank correlation metric.
These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
We regress them individually on each of the norm-based and PL-based metrics.

\input{table_3}

Results are summarized in Table~\ref{table:results} (and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} of the Supplementary Material).
For the mean, 
smaller RMSE, 
larger $R^2$, and 
larger Kendal-$\tau$ 
are desirable; 
and, for the standard deviation, smaller values are desirable.
Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
and PL-based metrics are much better for RMSE mean and standard deviation.
Model diagnostics (Supplementary Material) indicate many outliers and imperfect fits.
Overall, though, these and other results suggest our conclusions hold much more generally.