### 2021-02-18 submission/TOPML_Workshop/intro.tex
added into tex

We present our Theory of Heavy Tailed Self Regularization (HTSR) and
its application to the (overparameterized)  Deep Neural Networks (DNNs) 
to  predict the generalization accuracy and improve fine-tuning
of pretrained models by capturing the dominant correlation structure.
We analyze the correlations using techniques from 
Heavy Tailed Random Matrix Theory (HT-RMT)
and from  strongly correlated systems from theoretical
chemistry and  physics.  
We show how to predict trends in the reported test accuracies
across hundreds of pretrained DNNs without needing access
to training or test data by charactering the heavy tailed
structure of the empirical spectral densities (ESD) of 
the layer weight matrices.
In doing so, we show that different metrics, which
either depend on the shape or scale of the ESDs,
exhbit a Simpson's paradox when varying DNN model depth
vs regularization hyperparameters.
In particular,  the familar scale-based spectral norm
metric behave opposite theoretical predictions,
whereas our HT alpha-metric acts as predicted.
Moreover, the paradoxical behavior  can be resolved
by combining these shape and scale metrics with
our HT alpha-hat metric.
Finally,  And with access to training data, we show how to
smooth and capture the low rank structure to both accurately
predict the generalization errors and potentially improve
finetuning of pretrained DNNs.