### 2021-02-18 submission/TOPML_Workshop/paper.tex
paper on four pages, lots of fat removed

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2021} with \usepackage[nohyperref]{icml2021} above.
\usepackage{hyperref}

\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{overpic}
\usepackage{wrapfig}

\usepackage{siunitx}
\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}

% \DeclareMathOperator*{\cgrad}{grad}
\newcommand{\cgrad}{\operatorname{grad}}
\DeclareMathOperator*{\cdiv}{div}
\DeclareMathOperator*{\dd}{d}
\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\vect}{vec}
\DeclareMathOperator*{\real}{Re}

\newcommand{\code}[1]{\texttt{#1}}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}

\newcommand{\michael}[1]{{\color{red}\sf{[Michael: #1]}}}
\newcommand{\charles}[1]{{\color{blue}\sf{[Charles: #1]}}}

\newcommand{\fakesection}[1]{%
  \par\refstepcounter{section}% Increase section counter
  \sectionmark{#1}% Add section mark (header)
  \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}% Add section to ToC
  % Add more content here, if needed.
}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2021}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2021}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Using Heavy-Tailed Self-Regularization Theory}

\begin{document}

\twocolumn[
\icmltitle{Using Heavy-Tailed Self-Regularization Theory to Characterize Correlations in and Perform Diagnostics on State-of-the-art Overparameterized Models}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2021
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Charles H. Martin}{CC}
\icmlauthor{Michael W. Mahoney}{UCB}

\end{icmlauthorlist}

\icmlaffiliation{CC}{Department of Computer Science at Ben-Gurion University of the Negev, Israel.}
\icmlaffiliation{UCB}{ICSI and Department of Statistics at UC Berkeley, CA, USA.}

\icmlcorrespondingauthor{Omri Azencot}{azencot@cs.bgu.ac.il}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Differential Geometry, Sequence Models}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
% \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
Understanding state-of-the-art overparameterized neural networks requires detailed empirical studies as well as the development of predictive theory in light of those studies. 
Building on the theory of Heavy-Tailed Self-Regularization, we show that correlations in the weight matrices of state-of-the-art overparameterized neural networks are well-described by heavy-tailed random matrix theory.
Using the publicly-available WeightWatcher tool, we perform spectral analylsis on these weight matrices, and we show that fitted truncated power law exponents are strongly predictive of model quality, for publicly-available state-of-the-art pre-trained models, even in the absence of training or testing data.
We also show that these metrics can also be used to identify a Simpson's paradox in the data from a recent contest aimed at developing metrics to predice generalization in deep learning.
\end{abstract}

\section{Introduction}

\michael{Here is Charles core dump of intro; still to refine and incorporate with rest of text.}
We present our Theory of Heavy Tailed Self Regularization (HTSR) and
its application to the (overparameterized)  Deep Neural Networks (DNNs) 
to  predict the generalization accuracy and improve fine-tuning
of pretrained models by capturing the dominant correlation structure.
We analyze the correlations using techniques from 
Heavy Tailed Random Matrix Theory (HT-RMT)
and from  strongly correlated systems from theoretical
chemistry and  physics.  
We show how to predict trends in the reported test accuracies
across hundreds of pretrained DNNs without needing access
to training or test data by charactering the heavy tailed
structure of the empirical spectral densities (ESD) of 
the layer weight matrices.
In doing so, we show that different metrics, which
either depend on the shape or scale of the ESDs,
exhbit a Simpson's paradox when varying DNN model depth
vs regularization hyperparameters.
In particular,  the familar scale-based spectral norm
metric behave opposite theoretical predictions,
whereas our HT alpha-metric acts as predicted.
Moreover, the paradoxical behavior  can be resolved
by combining these shape and scale metrics with
our HT alpha-hat metric.
Finally,  And with access to training data, we show how to
smooth and capture the low rank structure to both accurately
predict the generalization errors and potentially improve
finetuning of pretrained DNNs.

\section{Statistical Mechanics and Heavy-Tailed Self Regularization}

%\citep{MM17_TR} is Charles: Rethinking generalization requires revisiting old ideas: statistical mechanics approaches and complex learning behavior''
%
%\citep{MM18_TR} is Charles: ``Implicit Self-Regularization in Deep Neural Networks: Evidence from Random Matrix Theory and Implications for Learning''
%
%\citep{MM19_HTSR_ICML} is Charles: ``Traditional and Heavy-Tailed Self Regularization in Neural Network Models''
%
%\citep{MM20_SDM} is Charles: ``Heavy-Tailed {U}niversality Predicts Trends in Test Accuracies for Very Large Pre-Trained Deep Neural Networks''
%
%\citep{MM19_KDD} is Charles: ``Statistical Mechanics Methods for Discovering Knowledge from Modern Production Quality Neural Networks''

Recently, Random Matrix Theory (RMT) was applied to analyze empirically the weight matrices of Deep Neural Networks (DNNs), including both production quality, pre-trained models such as AlexNet and Inception, and smaller models trained from scratch, such as LeNet5 and a miniature-AlexNet~\citep{MM18_TR} (see also \citep{MM17_TR, MM18_TR, MM19_HTSR_ICML, MM20_SDM, MM19_KDD}).
%%Empirical results clearly indicate that the empirical spectral density (ESD) of DNN layer matrices displays signatures of traditionally-regularized statistical models, even in the absence of exogenously specifying traditional forms of explicit regularization, such as dropout or weight norm~constraints.
%%
Building on these empirical results, 
%%and using relatively recent results in RMT, most notably its extension to Universality classes of heavy-tailed matrices, 
\citet{MM18_TR} then developed a phenomenological (statistical mechanics based, as opposed to statistical learning theory based) theory to identify ``5+1 Phases of Training,'' corresponding to increasing amounts of implicit self-regularization.
These phases can be observed during the training process as well as in the final learned DNNs.
For smaller and/or older DNNs, this implicit self-regularization is like traditional Tikhonov regularization, in that there is a ``size scale'' separating signal from noise.
For state-of-the-art (SOTA) DNNs, however, they identified a novel form of \emph{Heavy-Tailed Self-Regularization} (HT-ST).
This is similar to the self-organization seen in the statistical physics of disordered systems (such as classical models of actual neural activity); and 
it results from correlations arising at all size scales.
For DNNs, this arises due to the training process itself, when applied to very overparameterized models.

In the HT-SR Theory, one analyzes the eigenvalue spectrum, i.e., the Empirical Spectral Density (ESD), of the associated correlation matrices~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
From this, one characterizes the amount and form of correlation present in the DNN's weight matrices.
For each layer weight matrix $\mathbf{W}$, if we compute the eigenvalue spectrum of 
$ \mathbf{X} = \frac{1}{N}\mathbf{W}^{T}\mathbf{W} $
then HT-SR theory characterizes the correlations in a weight matrix by examining its ESD, $\rho(\lambda)$.
Empirically, the ESD can be well-fit to a truncated power law (PL) distribution, given as
$\rho(\lambda)\sim\lambda^{-\alpha} $,
which is (at least) valid within a bounded range of eigenvalues $\lambda\in[\lambda^{min},\lambda^{max}]$.

The original work on HT-SR Theory showed that for nearly every $\mathbf{W}$, i.e., among publicly-available SOTA models, the (bulk and tail) of the ESDs can be fit to a truncated PL, and that PL exponents $\alpha$ nearly all lie within the range $\alpha\in(1.5,5)$~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.
%
%%As for the mechanism responsible for these properties, statistical physics offers several possibilities~\citep{SornetteBook,nishimori01}, e.g., self organized criticality~\citep{SOC87,SOCat25yrs} or multiplicative noise in the stochastic optimization algorithms used to train these models~\citep{HodMah20A_TR,SorCon97}.
%%Alternatively, related techniques have been used to analyze correlations and information propogation in actual spiking neurons~\citep{SYYRP11,YKYP14}.
%%Our meta-analysis does not require knowledge of mechanisms; and it is not even clear that one mechanism is responsible for every case.
%
Crucially, HT-SR Theory predicts that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models.
%%The notion of ``size scale'' is well-defined in physical systems, to which this style of analysis is usually applied, but it is less well-defined in CV and NLP applications.
%%Informally, it would correspond to pixel groups that are at a greater distance in some metric, or between sentence parts that are at a greater distance in text.
%%Relatedly, previous work observed that smaller exponents $\alpha$ correspond to more implicit self-regularization and better generalization, and that we expect a linear correlation between $\hat{\alpha}$ and model quality~\citep{MM18_TR,MM19_HTSR_ICML,MM20_SDM}.

\section{Overall approach}

Test accuracies have been reported online for publicly-available pretrained pyTorch models~\citep{osmr}.
These models have been trained and evaluated on labeled data, using standard techniques.  
We do not have access to this data, and we have not trained any of the models ourselves. 
Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research, but uncommon in machine learning.
%
To evaluate the quality of SOTA DNNs, we consider the following metrics:
Frobenius Norm, $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$; 
Spectral Norm, $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$;
Weighted Alpha, $\hat{\alpha}=\alpha\log\lambda_{max}$; and
$\alpha$-Shatten Norm, $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. 
%\begin{eqnarray*}
%& & \text{Frobenius Norm: $\Vert\mathbf{W}\Vert^{2}_{F}=\Vert\mathbf{X}\Vert_{F}=\sum\nolimits_{i=1}^{M} \lambda_{i}$ } \\
%& & \text{Spectral Norm: $\Vert\mathbf{W}\Vert_{\infty}^{2}=\Vert\mathbf{X}\Vert_{\infty}=\lambda_{max}$ } \\
%& & \text{Weighted Alpha: $\hat{\alpha}=\alpha\log\lambda_{max}$ } \\
%& & \text{$\alpha$-Shatten Norm: $\Vert\mathbf{W}\Vert^{2\alpha}_{2\alpha}=\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}=\sum\nolimits_{i=1}^{M}\lambda_{i}^{\alpha}$. }
%\end{eqnarray*}
To perform diagnostics on potentially-problematic DNNs, we will decompose $\hat{\alpha}$ into its two components, $\alpha$ and $\lambda_{max}$.
One determines $\alpha$ for a given layer by fitting the ESD of that layer's weight matrix to a truncated PL, using the commonly accepted Maximum Likelihood method~\citep{CSN09_powerlaw,ABP14}.
Computations were performed with the publicly-available \texttt{WeightWatcher} tool (version 0.2.7)~\citep{weightwatcher_package}.

\section{Predicting Trends in the Quality of State-of-the-art Models}

%XXX.  ABSTRACT FROM \citep{MM20a_trends_TR}: ``Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data''

In many applications, one works with neural network models trained by someone else.
For such pretrained models, one may not have access to training data or test data.
Moreover, one may not know details about the model, e.g., the specifics of the training data, the loss function, the hyperparameter values, etc.
Given one or many pretrained models, it is a challenge to say anything about the expected performance or quality of the models.
We have applied HT-ST Theory to provide a detailed meta-analysis of hundreds of publicly-available pretrained models, examining both 
norm based capacity control metrics from statistical learning theory as well as PL based metrics from HT-ST theory.

To answer this question, we provide a detailed empirical analysis, evaluating quality metrics for pretrained DNN models, and we do so at scale.
Our approach may be viewed as a statistical meta-analysis of previously published work, where we consider a large suite of hundreds of publicly-available models, mostly from computer vision (CV) and natural language processing (NLP).
By now, there are many such state-of-the-art models that are publicly-available, e.g., 
hundreds of pretrained models in CV ($\ge 500$) and NLP ($\approx 100$).%
\footnote{When we began this work in 2018, there were fewer than tens of such models; now in 2020, there are hundreds of such models; and we expect that in a year or two there will be an order of magnitude or more of such models.}
For all these models, we have no access to training data or testing data, and we have no specific knowledge of the training/testing protocols.

We find that norm based metrics correlate well with reported test accuracies for well-trained models, but that they often cannot distinguish well-trained versus poorly-trained models.
We also find that power law based metrics can do much better---quantitatively better at discriminating among series of well-trained models with a given architecture; and qualitatively better at discriminating well-trained versus poorly-trained models.
These methods can be used to identify when a pretrained neural network has problems that cannot be detected simply by examining training/test accuracies.
%
%Here is a summary of our main results.
%First, norm-based metrics do a reasonably good job at predicting quality trends in well-trained CV/NLP models.
%Second, norm-based metrics may give spurious results when applied to poorly-trained models (e.g., models trained without enough data, etc.).
%For example, they may exhibit what we call Scale Collapse for these models. 
%Third, PL-based metrics can do much better at predicting quality trends in pretrained CV/NLP models.  
%In particular, 
%a weighted PL exponent (weighted by the log of the spectral norm of the corresponding layer) is 
%quantitatively better at discriminating among a series of well-trained versus very-well-trained models
%within a given architecture series; and
%the (unweighted) average PL exponent is 
%qualitatively better at discriminating well-trained versus poorly-trained models.
%Fourth, PL-based metrics can also be used to characterize fine-scale model properties, including what we call layer-wise Correlation Flow, in well-trained and poorly-trained models; and they can be used to evaluate model enhancements (e.g., distillation, fine-tuning, etc.).
%
Our work provides a theoretically-principled empirical evaluation---by far the largest, most detailed, and most comprehensive to date---and it uses HT-SR theory.%
\footnote{Performing such a meta-analysis of previously-published work is common in certain areas, but it is quite rare in ML, where the emphasis is on developing better training protocols.}

%\citep{BFT17_TR} is Bartlett: ``Spectrally-normalized margin bounds for neural networks''
%\citep{LMBx18_TR} is Poggio: ``A surprising linear relationship predicts test performance in deep networks''
There is related work \citep{BFT17_TR,LMBx18_TR}, but that work provides generalization bounds, modifies the models, requires access to data, etc., and so is quite different.

\section{A Simpson's Paradox---and the Need for Finer-scale Analysis}

XXX.  IF WANT A ONE SIZE FITS ALL METRIC, THEN CAN EXPECT A SIMPSONS PARADOX.  OUR TECHNIQUES CAN DIAGNOST THIS.  AND IN FACT WE SEE THAT.

\michael{Charles, put in a paragraph and the figure for the Simpson's paradox here.}

\charles{We apply our theory of Heavy Tailed Self-Regularization (HTSR) to predict the generalization accuracies of the pre-trained models provided in this contest, as a function of both model depth (L) and the solver hyperparameters (batch size, dropout, weight decay), and compare quality metrics that describe both the heavy-tailed shape and norm-based scale of the model layer weight matrices.  In doing so, we identify what amounts to a Simpson’s paradox, in which
shape and scale metrics show opposite behavior, and a combine metric resolves the paradox}

\begin{figure}[t!] 
    \centering
    \subfigure[Spectral Norm $\langle\log\Vert\mathbf{W}\Vert_{\infty}\rangle$]{
      \includegraphics[width=4.0cm]{img/simpsons-task2-logspectralnorm.png}
    }
    \subfigure[Alpha Hat $\hat{\alpha}$]{
     \includegraphics[width=4cm]{img/simpsons-task2-alpha_weighted.png}
    }
    \caption{Test accuracy as a function of the
Log Spectral Norm $\log\Vert\mathbf{W}\Vert_{\infty}$ 
and the HTSR Alpha Hat $\hat{\alpha}$ metrics
as model hyperparameters and depth are varied. 
Models from the NeurIPS Predicting Generalization contest.}
    \label{fig:simpsons}
\end{figure}

\citep{JNBx19_fantastic_TR} is the terrible fantastic paper: ``Fantastic Generalization Measures and Where to Find Them''

\citep{JFYx20_contest_v10} is the contest: ``{NeurIPS} 2020 Competition: Predicting Generalization in Deep Learning (Version 1.0)''

\citep{BHO75} is Bickel: ``Sex Bias in Graduate Admissions: Data from Berkeley''

\section{Conclusion}

\michael{I may cut this stuff.}
We have shown that weight matrix analysis derived from HT-SR Theory can be used to predict trends in the quality of state-of-the-art neural networks without access to training or testing data as well as identify Simpson's-like paradoxes in publicly-available NN models.
The HT-SR Theory underlying our analysis is quite different than theoretical approaches popular in ML, and indeed it is quite different than typical statistical mechanics approaches to NNs \citep{BKPx20}.
Recently, however, work motivated by HT-SR Theory has 
demonstrated that multiplicative noise in stochastic optimization leads to heavy tails~\citep{HodMah20A_TR,GSZ20_TR}; 
provided a random matrix analysis of random {F}ourier features, going beyond the {G}aussian kernel, to provide a precise phase transition and illustrate the corresponding double descent~\citep{ZCM20_TR}; 
provided exact expressions for double descent and implicit regularization via a notion of surrogate random design \citep{DLM19_Exact_TR}; 
demonstrated that good classifiers are abundant in the interpolating regime \citep{TKM20_TR};
shown how injecting noise into hidden states of recurrent NNs leads to implicit regularization~\citep{LEHM21_TR,Mah12}; and 
shown that analogous phase transitions hold for the {C}olumn {S}ubset {S}election and the {N}ystrom method~\citep{DKM20_TR}. 
We expect that the empirical results identified by tools from HT-SR Theory will provide guidance and constraints for future development of overparameterized machine learning models.

\bibliographystyle{icml2021}
\bibliography{dnns}

\end{document}